Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-12 paper/agree.tex
Add text on synthesis

Add text to this section that defines the meaning of the contract over streams. State the prove guarantee from AGREE is terms of the streams. The statement will be used in the synthesis section to argue correctness.

>>>>

\emph{Add text to this section that defines the meaning of the contract over streams. State the prove guarantee from AGREE is terms of the streams. The statement will be used in the synthesis section to argue correctness.}

### 2021-05-12 paper/example.tex
Add text on synthesis

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. It raises an alert if ever the specified temporal properties are violated. The auto-generated AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and that response must come with the request or in the next step after the request. As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.

@@

The high-assurance components are automatically synthesized from the AGREE specifications to equivalent models in the CakeML language. CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.

A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML. A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML. In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.

>>>>

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. It raises an alert if ever the specified temporal properties are violated. The AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and, also,  that response must come with the request or in the next step after the request. As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.

@@

The high-assurance components are automatically synthesized from the AGREE specifications to equivalent models in the CakeML language. CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.

A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML. A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the generated code. In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.

### 2021-05-12 paper/intro.tex
Add text on synthesis

A novel aspect of the approach is the ability to automatically transformation the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities. The behavior of the model is given by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result of the cyber-vulnerability analysis to assume properties on the input and and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to implement the hardening, AGREE fails to prove the system cyber-hardened.

The two automatic transforms discussed in this paper are the insertion of filters to prevent malformed data from malicious actors from being propagated to downstream components, and monitors to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML. This paper agues that synthesis to CakeML to be correct. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system in that in detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

>>>>

A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities. The behavior of the model is given by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result of the cyber-vulnerability analysis to assume properties on the input and and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to implement the hardening, AGREE fails to prove the system cyber-hardened.

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

### 2021-05-12 paper/synthesis.tex
Add text on synthesis

A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs. The computation for the component in this work is defined entirely by a step function:
\[
\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
\]
The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
\[
\begin{array}{ll}
 \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
 (v_1,\ldots) & = \konst{readState}() ; \\
 ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
\end{array}
\]
It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.

\newsavebox{\cml}
\begin{lrbox}{\cml}
\begin{lstlisting}[style=myML]
fun filter_step () =
let val () = Utils.clear_buf buffer
    val () = API.callFFI "get_input" "" buffer
    val string = Utils.buf2string buffer
    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
in
    if policy
    then
        let val string = Word8Array.substring buffer 1
                          (Word8Array.length buffer - 1)
        in API.callFFI "put_output" string Utils.emptybuf
        end
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\cml}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:cakeml}
\end{figure}

The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
The auto-generated AGREE specification raises an alert output when the relation is violated.

>>>>

Synthesis maps from model and specifications to code. The synthesis algorithm traverses the system architecture looking
for occurrences of filter and monitor specifications. For each such 
occurrence it generates a CakeML program. In the following, we examine both filter and monitor synthesis. The latter is typically much more involved, and we will devote more attention to it.

\subsection{Filter Generation}

A filter is intended to be very simple; it is expected to have one
input port and one output; messages on the input that the filter
policy admits pass unchanged to the output port; all others are not
passed on. We have explored in our work two kinds of filter. In the
first, a relatively shallow scan of the input buffer can enforce the
policy. For example, we have used the expressive power of contiguity
types \cite{contig} to enforce \emph{lightweight} bounds constraints
on GPS coordinates in UxAS messages. On the other hand, the second
kind of filter will parse the input buffer into a data structure
specified in AGREE and apply an user-defined \emph{well-formedness}
property, also specified in AGREE, to the data. This allows
arbitrarily complex well-formedness checking.

The decision of a filter is made and performed within one thread
invocation. Thus, in its given time slice, the filter does the following:

\begin{enumerate}

\item checks to see if there is any input available; if there is none
then it yields control; otherwise,

\item the input is read (and parsed if need be),

\item the wellformedness predicate is evaluated,

\item if the predicate returns \konst{true} then the input buffer
 is copied to the output; otherwise no action is taken, and

\item the filter yields control
\end{enumerate}

\begin{remark}[Partiality]

The role of partiality should be emphasized: steps 2 and 3 can fail;
the data might not be parseable or the wellformedness computation
could be badly written and fail at run time. In such cases, the filter
should recover and yield control without passing the input onwards. In
these cases, the filter is behaving as it should, but there are also
cases in which a correctly specified filter would not meet its
specification at runtime. This situation arises when the
filter \emph{ought} to accept a message, but lack of resources result
in the filter failing to do so. Examples of this would be, for
example, if the parse of a message needed more space than allocated;
another example would be if the timeslice provided by the scheduler is
too small for the wellformedness computation to finish.

\end{remark}

\emph{Need some discussion of filters and their step-wise propertie s in relation to their infinitary properties. Reference to Johannes' work.}

\newsavebox{\cml}
\begin{lrbox}{\cml}
\begin{lstlisting}[style=myML]
fun filter_step () =
let val () = Utils.clear_buf buffer
    val () = API.callFFI "get_input" "" buffer
in
    if WELL_FORMED_AUTOMATION_RESPONSE buffer
    then
      API.callFFI "put_output" buffer Utils.emptybuf
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\cml}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:filter-cakeml}
\end{figure}

The synthesized CakeML code for the filter is shown in \figref{fig:filter-cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
The auto-generated AGREE specification raises an alert output when the relation is violated.

% A \emph{system} is a collection of \emph{components}, \emph{connections}
% between components, a \emph{scheduler} to order execution, and a
% \emph{system environment} for primary inputs.

\subsection{Monitor Generation}

Monitors, since they are intended to track and analyze the externally visible behavior of system components through time, require more computational features than filters. In particular, our conception of a monitor is a predicate on its input (and output) streams, being able to access the value of a stream at any earlier point in time,  if necessary. Thus monitors commonly use state to keep track of earlier values, unlike filters which are stateless. This reasoning leads us to specify the  computation for a monitor component  by a step function of the following form:
\[
\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
\]
Thus, in its given timeslice, a monitor evaluates the \konst{stepFn} on its inputs and the current values of the state variables. In detail, it takes the following steps:

\begin{enumerate}

\item each available input is parsed into data of the type specified
by the port type;

\item the stateful variables in $\mathit{stateVars}$  are evaluated in dependency order.

\item values of outputs are computed

\item outputs are written and the new state is written

\item control is yielded
\end{enumerate}

Our earlier remarks on partiality apply here too, of course.

The scheduler \emph{activates} components in some order.It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
 Activating a monitor component takes the form of the following pseudo-code: 
\[
\begin{array}{ll}
 \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
 (v_1,\ldots) & = \konst{readState}() ; \\
 ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
\end{array}
\]

\paragraph{Initialization}
A monitor may need to make a certain minimum number of observations before being able to make a meaningful assessment of behavior. Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization}. In order for correct code to be generated, monitor specifications need to spell out the values of output ports when in such initialization phases. For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at time $t$ and $t+1$ is such that the distance between the two locations is unusually large. Such a monitor needs two measurements before making its first judgement, but at the time of its first output, only one measurement will have been made. The specification for this monitor has to explicitly say what the correct first output is.
